---
format: gfm
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RUtilpol

<!-- badges: start -->
[![CRAN-status](https://www.r-pkg.org/badges/version/RUtilpol)](https://CRAN.R-project.org/package=RUtilpol)
[![R-CMD-check](https://github.com/HOPE-UIB-BIO/R-Utilpol-package/workflows/R-CMD-check/badge.svg)](https://github.com/HOPE-UIB-BIO/R-Utilpol-package/actions)
<!-- badges: end -->

R-Utilpol is an R package proving utility functions for package development.

## WARNING

:bangbang: This package is under active development. Therefore, the package is subject to change. :bangbang:

## Installation

You can install the development version of RUtilpol from [GitHub](https://github.com/) with:

```{r install, eval = FALSE}
# install.packages("remotes")
remotes::install_github("HOPE-UIB-BIO/R-Utilpol-package")
```

## Usage

The main purpose of {RUtilpol} is to provide utility (helper) functions. All functions can be grouped into 3 main categories:

1. **Package development** - Function used within other packages (e.g., `RUtilpol::check_class()` for safety check, `RUtilpol::output_warning()` - for console message, `RUtilpol::replace_null_with_na()` for data wrangling ,...).

2. [**File management**](#file-management) - Functions for saving and loading functions for better handling of files in large projects (see examples below).

3. [**Extraction of spatial information**](#extraction-of-spatial-information) - Functions to assign information to points based on a raster of shapefile (see examples below).

### File management

```{r}

```

### Extraction of spatial information

```{r}

```